---
title: "`scLANE` Simulation Study - Simulated Data Quality Control"
subtitle: "UF Dept. of Biostatistics - Bacher Group"
author: "Jack Leary" 
date: "`r Sys.Date()`"
output:
  html_document:
    theme: journal
    highlight: tango
    code_folding: show
    code_download: true 
    toc: true 
    toc_depth: 2
    toc_float: true
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE); set.seed(312)  # lucky seed
```

# Libraries 

```{r, results='hide', message=FALSE, warning=FALSE}
library(dplyr)                 # data manipulation 
library(scran)                 # scRNA tools 
library(scater)                # more scRNA tools
library(Seurat)                # scRNA methods & data structures
library(ggplot2)               # plots
library(targets)               # pipeline tools
library(paletteer)             # plot colors 
library(patchwork)             # plot combination
library(SingleCellExperiment)  # scRNA data structures 
```

# Color Palette

We'll be doing some complex visualization, so we'll pick out a good color palette beforehand to make our code more readable & ensure that we don't run in to errors caused by the palette not being long enough. 

```{r, message=FALSE, warning=FALSE}
palette_cluster <- paletteer_d("ggsci::default_igv")
```

# Brain Reference 

First we'll load in the brain reference dataset from the `{scRNAseq}` package & process it like we would our simulated datasets. Please ignore the calls to `gc()` littered throughout the code, it was so difficult to get all these datasets into memory & processed without my R session crashing over & over again. 

```{r, results='hide', message=FALSE, warning=FALSE}
brain_data <- scRNAseq::LaMannoBrainData(which = "human-embryo")
brain_data_clean <- brain_data[rowSums(counts(brain_data) > 0) >= 3, ]
brain_data_clean <- logNormCounts(brain_data_clean)
var_decomp <- modelGeneVar(brain_data_clean)
top2k_hvgs <- getTopHVGs(var_decomp, n = 2000)
brain_data_clean <- runPCA(brain_data_clean, subset_row = top2k_hvgs)
reducedDim(brain_data_clean, "PCAsub") <- reducedDim(brain_data_clean, "PCA")[, 1:30, drop = FALSE]
brain_data_clean <- runUMAP(brain_data_clean, 
                            dimred = "PCAsub", 
                            n_dimred = 1:30)
g <- buildSNNGraph(brain_data_clean, 
                   use.dimred = "PCAsub", 
                   k = 30)
clusters <- igraph::cluster_louvain(graph = g)$membership
colLabels(brain_data_clean) <- factor(clusters)
brain_data_clean <- as.Seurat(brain_data_clean, 
                              counts = "counts", 
                              data = "logcounts")
gc(full = TRUE)
```

Next we create a table of summary statistics & dataset characteristics. We then create a `{ggplot2}`-friendly version of the table using `{gridExtra}` & `{gtable}`. This allows us to include the table as part of a plot object. 

```{r, results='hide', message=FALSE, warning=FALSE}
n_cells <- ncol(brain_data_clean)
n_genes <- nrow(brain_data_clean)
sparsity_count <- mean(brain_data_clean@assays$originalexp@counts == 0)
mean_count <- mean(brain_data_clean@assays$originalexp@counts)
med_count <- ifelse(sparsity_count > 0.5, 0, median(brain_data_clean@assays$originalexp@counts))
sd_count <- sd(brain_data_clean@assays$originalexp@counts)
var_count <- sd_count^2
range_count <- range(brain_data_clean@assays$originalexp@counts)
summary_df <- data.frame(metric = c("Mean", "Median", "S.D.", "Variance", "Range", "Sparsity"), 
                         value = c(round(mean_count, 2), 
                                   med_count, 
                                   round(sd_count, 2), 
                                   round(var_count, 2),
                                   paste0("(", range_count[1], ", ", range_count[2], ")"), 
                                   paste0(round(sparsity_count, 4) * 100, "%")))
plot_table <- gridExtra::tableGrob(summary_df, 
                                   rows = NULL, 
                                   cols = c("Metric", "Value"), 
                                   theme = gridExtra::ttheme_minimal(core = list(fg_params = list(hjust = 0, x = 0.05)), 
                                                                     colhead = list(fg_params = list(hjust = 0, x = 0.05)))) %>% 
              gtable::gtable_add_grob(grobs = grid::rectGrob(gp = grid::gpar(fill = NA, lwd = 2)), 
                                      t = 2, 
                                      b = nrow(.), 
                                      l = 1, 
                                      r = ncol(.)) %>% 
              gtable::gtable_add_grob(grobs = grid::rectGrob(gp = grid::gpar(fill = NA, lwd = 2)), 
                                      t = 1, 
                                      l = 1, 
                                      r = ncol(.))
gc(full = TRUE)
```

The first plots we want are histograms of the raw & normalized counts.

```{r, message=FALSE, warning=FALSE}
p0 <- data.frame(x = as.numeric(brain_data_clean@assays$originalexp@counts)) %>% 
      ggplot(aes(x = x)) + 
      geom_histogram(fill = "deepskyblue3", 
                     color = "dodgerblue4", 
                     alpha = 0.75, 
                     linewidth = 0.75) + 
      scale_y_continuous(labels = scales::label_scientific()) + 
      scale_x_continuous(labels = scales::label_comma()) + 
      labs(x = "Raw Expression", y = "Frequency") + 
      theme_classic(base_size = 14)
p1 <- data.frame(x = as.numeric(brain_data_clean@assays$originalexp@data)) %>% 
      ggplot(aes(x = x)) + 
      geom_histogram(fill = "darkseagreen", 
                     color = "forestgreen", 
                     alpha = 0.75, 
                     linewidth = 0.75) + 
      scale_y_continuous(labels = scales::label_scientific()) + 
      labs(x = "Normalized Expression", y = "Frequency") + 
      theme_classic(base_size = 14)
```

We next make a UMAP & PCA plot of the unsupervised clustering. 

```{r, message=FALSE, warning=FALSE}
p2 <- data.frame(UMAP1 = brain_data_clean@reductions$UMAP@cell.embeddings[, 1], 
                 UMAP2 = brain_data_clean@reductions$UMAP@cell.embeddings[, 2], 
                 cluster = brain_data_clean$label) %>% 
      ggplot(aes(x = UMAP1, y = UMAP2, color = cluster)) + 
      geom_point() + 
      scale_color_manual(values = palette_cluster) + 
      labs(x = "UMAP 1", 
           y = "UMAP 2", 
           color = "Louvain Cluster") + 
      theme_classic(base_size = 14) + 
      theme(axis.text = element_blank(), 
            axis.ticks = element_blank()) + 
      guides(color = guide_legend(override.aes = list(size = 4)))
p3 <- data.frame(PC1 = brain_data_clean@reductions$PCA@cell.embeddings[, 1], 
                 PC2 = brain_data_clean@reductions$PCA@cell.embeddings[, 2], 
                 cluster = brain_data_clean$label) %>% 
      ggplot(aes(x = PC1, y = PC2, color = cluster)) + 
      geom_point() + 
      scale_color_manual(values = palette_cluster) + 
      labs(x = "PC 1", 
           y = "PC 2", 
           color = "Louvain Cluster") + 
      theme_classic(base_size = 14) + 
      theme(legend.position = "none", 
            axis.text = element_blank(), 
            axis.ticks = element_blank())
```

We align everything using the `{patchwork}` package, save the figure, and plot it all. 

```{r, fig.width=13, fig.height=8, message=FALSE, warning=FALSE}
p4a <- (p0 | p1) / (p2 | p3) + 
       plot_layout(guides = "collect")
p4b <- (p4a | plot_table) + 
       plot_layout(ncol = 2, widths = c(3, 1)) + 
       plot_annotation(title = paste0("Metrics for Brain Reference Dataset"), 
                       theme = theme(plot.title = element_text(face = "bold", size = 16)))
ggsave(filename = "QC_brain_reference.pdf",
       plot = p4b, 
       device = "pdf", 
       path = "/blue/rbacher/j.leary/repos/scLANE-Simulations/Figures/Sim_QC/", 
       width = 13,
       height = 8, 
       units = "in", 
       dpi = "retina")
p4b
```

## Single-subject

We load the single-subject simulated datasets into a list & turn them into `{Seurat}` objects for plotting. 

```{r, results='hide', message=FALSE, warning=FALSE}
# 100 cells 
tar_load(sim_single_brain_1_100, store = "../store_simulation/")
tar_load(sim_single_brain_5_100, store = "../store_simulation/")
tar_load(sim_single_brain_10_100, store = "../store_simulation/")
tar_load(sim_single_brain_20_100, store = "../store_simulation/")
# 500 cells 
tar_load(sim_single_brain_1_500, store = "../store_simulation/")
tar_load(sim_single_brain_5_500, store = "../store_simulation/")
tar_load(sim_single_brain_10_500, store = "../store_simulation/")
tar_load(sim_single_brain_20_500, store = "../store_simulation/")
# 1,000 cells 
tar_load(sim_single_brain_1_1000, store = "../store_simulation/")
tar_load(sim_single_brain_5_1000, store = "../store_simulation/")
tar_load(sim_single_brain_10_1000, store = "../store_simulation/")
tar_load(sim_single_brain_20_1000, store = "../store_simulation/")
# 2,500 cells
tar_load(sim_single_brain_1_2500, store = "../store_simulation/")
tar_load(sim_single_brain_5_2500, store = "../store_simulation/")
tar_load(sim_single_brain_10_2500, store = "../store_simulation/")
tar_load(sim_single_brain_20_2500, store = "../store_simulation/")
# 5,000 cells 
tar_load(sim_single_brain_1_5000, store = "../store_simulation/")
tar_load(sim_single_brain_5_5000, store = "../store_simulation/")
tar_load(sim_single_brain_10_5000, store = "../store_simulation/")
tar_load(sim_single_brain_20_5000, store = "../store_simulation/")
# coerce to list & process
obj_list <- purrr::map(ls(pattern = "sim_single_brain"), \(sim) {
  obj <- eval(as.symbol(sim))
  reducedDim(obj, "PCAsub") <- NULL
  obj <- as.Seurat(obj, 
                   counts = "counts", 
                   data = "logcounts")
  obj@meta.data <- mutate(obj@meta.data, 
                          perc_deg = paste0(as.numeric(gsub("_.*", "", gsub("sim_single_brain_", "", sim))), "%"), 
                          n_cells = as.character(ncol(obj)), 
                          n_genes = as.character(nrow(obj)), 
                          sce_name = sim)
  return(obj)
})
rm(list = ls(pattern = "sim_single_brain")); gc(full = TRUE)
```

```{r}
test <- run_scLANE_GLM(sim_single_brain_5_500, n.genes.sample = 200)
```


Iterating over the datasets, we print each QC plot & save them to PDFs. 

```{r, fig.width=13, fig.height=8, results='hold', message=FALSE, warning=FALSE}
purrr::walk(obj_list, function(z) {
  # gather metadata 
  obj_name <- z@meta.data$sce_name[1]
  n_cells <- z@meta.data$n_cells[1]
  n_genes <- z@meta.data$n_genes[1]
  perc_deg <- z@meta.data$perc_deg[1]
  # summary stat table 
  sparsity_count <- mean(z@assays$originalexp@counts == 0)
  mean_count <- mean(z@assays$originalexp@counts)
  med_count <- ifelse(sparsity_count > 0.5, 0, median(z@assays$originalexp@counts))
  sd_count <- sd(z@assays$originalexp@counts)
  var_count <- sd_count^2 
  range_count <- range(z@assays$originalexp@counts)
  summary_df <- data.frame(metric = c("Mean", "Median", "S.D.", "Variance", "Range", "Sparsity"), 
                           value = c(round(mean_count, 2), 
                                     round(med_count, 2), 
                                     round(sd_count, 2), 
                                     round(var_count, 2),
                                     paste0("(", range_count[1], ", ", range_count[2], ")"), 
                                     paste0(round(sparsity_count, 4) * 100, "%")))
  # create counts histogram
  p0 <- data.frame(x = as.numeric(z@assays$originalexp@counts)) %>% 
      ggplot(aes(x = x)) + 
      geom_histogram(fill = "deepskyblue3", 
                     color = "dodgerblue4", 
                     alpha = 0.75, 
                     linewidth = 0.75) + 
      scale_y_continuous(labels = scales::label_scientific()) + 
      scale_x_continuous(labels = scales::label_comma()) + 
      labs(x = "Raw Expression", y = "Frequency") + 
      theme_classic(base_size = 14)
  # create log counts histogram
  p1 <- data.frame(x = as.numeric(z@assays$originalexp@data)) %>% 
        ggplot(aes(x = x)) + 
        geom_histogram(fill = "darkseagreen", 
                       color = "forestgreen", 
                       alpha = 0.75, 
                       linewidth = 0.75) + 
        scale_y_continuous(labels = scales::label_scientific()) + 
        labs(x = "Normalized Expression", y = "Frequency") + 
        theme_classic(base_size = 14)
  # create UMAP by cluster 
  p2 <- data.frame(UMAP1 = z@reductions$UMAP@cell.embeddings[, 1], 
                   UMAP2 = z@reductions$UMAP@cell.embeddings[, 2], 
                   cluster = z$label) %>% 
        ggplot(aes(x = UMAP1, y = UMAP2, color = cluster)) + 
        geom_point() + 
        scale_color_manual(values = palette_cluster) + 
        labs(x = "UMAP 1",
             y = "UMAP 2", 
             color = "Louvain Cluster") + 
        theme_classic(base_size = 14) + 
        theme(axis.text = element_blank(), 
              axis.ticks = element_blank()) + 
        guides(color = guide_legend(override.aes = list(size = 4)))
  # create PCA of cell ordering 
  p3 <- data.frame(PC1 = z@reductions$PCA@cell.embeddings[, 1], 
                   PC2 = z@reductions$PCA@cell.embeddings[, 2], 
                   cell_time = z$cell_time_normed) %>% 
        ggplot(aes(x = PC1, y = PC2, color = cell_time)) + 
        geom_point() + 
        scale_color_gradientn(colors = paletteer_d("wesanderson::Zissou1")) + 
        labs(x = "PC 1", 
             y = "PC 2", 
             color = "True Ordering") + 
        theme_classic(base_size = 14) + 
        theme(axis.text = element_blank(), 
              axis.ticks = element_blank())
  # table of simulation parameters 
  param_df <- data.frame(metric = c("Number of Cells", "Number of Genes", "% Dynamic Genes"), 
                         value = c(as.character(n_cells), as.character(n_genes), perc_deg))
  plot_table <- rbind(summary_df, param_df) %>% 
                gridExtra::tableGrob(rows = NULL, 
                                     cols = c("Metric", "Value"), 
                                     theme = gridExtra::ttheme_minimal(core = list(fg_params = list(hjust = 0, x = 0.05)), 
                                                                       colhead = list(fg_params = list(hjust = 0, x = 0.05)))) %>% 
                gtable::gtable_add_grob(grobs = grid::rectGrob(gp = grid::gpar(fill = NA, lwd = 2)), 
                                        t = 2, 
                                        b = nrow(.), 
                                        l = 1, 
                                        r = ncol(.)) %>% 
                gtable::gtable_add_grob(grobs = grid::rectGrob(gp = grid::gpar(fill = NA, lwd = 2)), 
                                        t = 1, 
                                        l = 1, 
                                        r = ncol(.))
  # align everything 
  p4a <- (p0 | p1) / (p2 | p3) + 
         plot_layout(guides = "collect")
  p4b <- (p4a | plot_table) + 
         plot_layout(ncol = 2, widths = c(3, 1)) + 
         plot_annotation(title = paste0("Metrics for dataset: ", obj_name), 
                         theme = theme(plot.title = element_text(face = "bold", size = 16)))
  # save & print plot
  ggsave(filename = paste0("QC_", obj_name, ".pdf"),
         plot = p4b, 
         device = "pdf", 
         path = "/blue/rbacher/j.leary/repos/scLANE-Simulations/Figures/Sim_QC/", 
         width = 13,
         height = 8, 
         units = "in", 
         dpi = "retina")
  print(p4b)
  # cleanup 
  sink(tempfile())
  rm(p0, p1, p2, p3, p4a, p4b, plot_table); gc(full = TRUE)
  sink()
})
rm(obj_list)
```

# Pancreatic Endocrinogenesis Reference 

First we'll load in the brain reference dataset from the `{scRNAseq}` package & process it like we would our simulated datasets. Please ignore the calls to `gc()` littered throughout the code, it was so difficult to get all these datasets into memory & processed without my R session crashing over & over again. 

```{r, results='hide', message=FALSE, warning=FALSE}
brain_data <- scRNAseq::LaMannoBrainData(which = "human-embryo")
brain_data_clean <- brain_data[rowSums(counts(brain_data) > 0) >= 3, ]
brain_data_clean <- logNormCounts(brain_data_clean)
var_decomp <- modelGeneVar(brain_data_clean)
top2k_hvgs <- getTopHVGs(var_decomp, n = 2000)
brain_data_clean <- runPCA(brain_data_clean, subset_row = top2k_hvgs)
reducedDim(brain_data_clean, "PCAsub") <- reducedDim(brain_data_clean, "PCA")[, 1:30, drop = FALSE]
brain_data_clean <- runUMAP(brain_data_clean, 
                            dimred = "PCAsub", 
                            n_dimred = 1:30)
g <- buildSNNGraph(brain_data_clean, 
                   use.dimred = "PCAsub", 
                   k = 30)
clusters <- igraph::cluster_louvain(graph = g)$membership
colLabels(brain_data_clean) <- factor(clusters)
brain_data_clean <- as.Seurat(brain_data_clean, 
                              counts = "counts", 
                              data = "logcounts")
gc(full = TRUE)
```

Next we create a table of summary statistics & dataset characteristics. We then create a `{ggplot2}`-friendly version of the table using `{gridExtra}` & `{gtable}`. This allows us to include the table as part of a plot object. 

```{r, results='hide', message=FALSE, warning=FALSE}
n_cells <- ncol(brain_data_clean)
n_genes <- nrow(brain_data_clean)
sparsity_count <- mean(brain_data_clean@assays$originalexp@counts == 0)
mean_count <- mean(brain_data_clean@assays$originalexp@counts)
med_count <- ifelse(sparsity_count > 0.5, 0, median(brain_data_clean@assays$originalexp@counts))
sd_count <- sd(brain_data_clean@assays$originalexp@counts)
var_count <- sd_count^2
range_count <- range(brain_data_clean@assays$originalexp@counts)
summary_df <- data.frame(metric = c("Mean", "Median", "S.D.", "Variance", "Range", "Sparsity"), 
                         value = c(round(mean_count, 2), 
                                   med_count, 
                                   round(sd_count, 2), 
                                   round(var_count, 2),
                                   paste0("(", range_count[1], ", ", range_count[2], ")"), 
                                   paste0(round(sparsity_count, 4) * 100, "%")))
plot_table <- gridExtra::tableGrob(summary_df, 
                                   rows = NULL, 
                                   cols = c("Metric", "Value"), 
                                   theme = gridExtra::ttheme_minimal(core = list(fg_params = list(hjust = 0, x = 0.05)), 
                                                                     colhead = list(fg_params = list(hjust = 0, x = 0.05)))) %>% 
              gtable::gtable_add_grob(grobs = grid::rectGrob(gp = grid::gpar(fill = NA, lwd = 2)), 
                                      t = 2, 
                                      b = nrow(.), 
                                      l = 1, 
                                      r = ncol(.)) %>% 
              gtable::gtable_add_grob(grobs = grid::rectGrob(gp = grid::gpar(fill = NA, lwd = 2)), 
                                      t = 1, 
                                      l = 1, 
                                      r = ncol(.))
gc(full = TRUE)
```

The first plots we want are histograms of the raw & normalized counts.

```{r, message=FALSE, warning=FALSE}
p0 <- data.frame(x = as.numeric(brain_data_clean@assays$originalexp@counts)) %>% 
      ggplot(aes(x = x)) + 
      geom_histogram(fill = "deepskyblue3", 
                     color = "dodgerblue4", 
                     alpha = 0.75, 
                     linewidth = 0.75) + 
      scale_y_continuous(labels = scales::label_scientific()) + 
      scale_x_continuous(labels = scales::label_comma()) + 
      labs(x = "Raw Expression", y = "Frequency") + 
      theme_classic(base_size = 14)
p1 <- data.frame(x = as.numeric(brain_data_clean@assays$originalexp@data)) %>% 
      ggplot(aes(x = x)) + 
      geom_histogram(fill = "darkseagreen", 
                     color = "forestgreen", 
                     alpha = 0.75, 
                     linewidth = 0.75) + 
      scale_y_continuous(labels = scales::label_scientific()) + 
      labs(x = "Normalized Expression", y = "Frequency") + 
      theme_classic(base_size = 14)
```

We next make a UMAP & PCA plot of the unsupervised clustering. 

```{r, message=FALSE, warning=FALSE}
p2 <- data.frame(UMAP1 = brain_data_clean@reductions$UMAP@cell.embeddings[, 1], 
                 UMAP2 = brain_data_clean@reductions$UMAP@cell.embeddings[, 2], 
                 cluster = brain_data_clean$label) %>% 
      ggplot(aes(x = UMAP1, y = UMAP2, color = cluster)) + 
      geom_point() + 
      scale_color_manual(values = palette_cluster) + 
      labs(x = "UMAP 1", 
           y = "UMAP 2", 
           color = "Louvain Cluster") + 
      theme_classic(base_size = 14) + 
      theme(axis.text = element_blank(), 
            axis.ticks = element_blank()) + 
      guides(color = guide_legend(override.aes = list(size = 4)))
p3 <- data.frame(PC1 = brain_data_clean@reductions$PCA@cell.embeddings[, 1], 
                 PC2 = brain_data_clean@reductions$PCA@cell.embeddings[, 2], 
                 cluster = brain_data_clean$label) %>% 
      ggplot(aes(x = PC1, y = PC2, color = cluster)) + 
      geom_point() + 
      scale_color_manual(values = palette_cluster) + 
      labs(x = "PC 1", 
           y = "PC 2", 
           color = "Louvain Cluster") + 
      theme_classic(base_size = 14) + 
      theme(legend.position = "none", 
            axis.text = element_blank(), 
            axis.ticks = element_blank())
```

We align everything using the `{patchwork}` package, save the figure, and plot it all. 

```{r, fig.width=13, fig.height=8, message=FALSE, warning=FALSE}
p4a <- (p0 | p1) / (p2 | p3) + 
       plot_layout(guides = "collect")
p4b <- (p4a | plot_table) + 
       plot_layout(ncol = 2, widths = c(3, 1)) + 
       plot_annotation(title = paste0("Metrics for Brain Reference Dataset"), 
                       theme = theme(plot.title = element_text(face = "bold", size = 16)))
ggsave(filename = "QC_brain_reference.pdf",
       plot = p4b, 
       device = "pdf", 
       path = "/blue/rbacher/j.leary/repos/scLANE-Simulations/Figures/Sim_QC/", 
       width = 13,
       height = 8, 
       units = "in", 
       dpi = "retina")
p4b
```

## Single-subject

We load the single-subject simulated datasets into a list & turn them into `{Seurat}` objects for plotting. 

```{r, results='hide', message=FALSE, warning=FALSE}
# 100 cells 
tar_load(sim_single_brain_1_100, store = "../store_simulation/")
tar_load(sim_single_brain_5_100, store = "../store_simulation/")
tar_load(sim_single_brain_10_100, store = "../store_simulation/")
tar_load(sim_single_brain_20_100, store = "../store_simulation/")
# 500 cells 
tar_load(sim_single_brain_1_500, store = "../store_simulation/")
tar_load(sim_single_brain_5_500, store = "../store_simulation/")
tar_load(sim_single_brain_10_500, store = "../store_simulation/")
tar_load(sim_single_brain_20_500, store = "../store_simulation/")
# 1,000 cells 
tar_load(sim_single_brain_1_1000, store = "../store_simulation/")
tar_load(sim_single_brain_5_1000, store = "../store_simulation/")
tar_load(sim_single_brain_10_1000, store = "../store_simulation/")
tar_load(sim_single_brain_20_1000, store = "../store_simulation/")
# 2,500 cells
tar_load(sim_single_brain_1_2500, store = "../store_simulation/")
tar_load(sim_single_brain_5_2500, store = "../store_simulation/")
tar_load(sim_single_brain_10_2500, store = "../store_simulation/")
tar_load(sim_single_brain_20_2500, store = "../store_simulation/")
# 5,000 cells 
tar_load(sim_single_brain_1_5000, store = "../store_simulation/")
tar_load(sim_single_brain_5_5000, store = "../store_simulation/")
tar_load(sim_single_brain_10_5000, store = "../store_simulation/")
tar_load(sim_single_brain_20_5000, store = "../store_simulation/")
# coerce to list & process
obj_list <- purrr::map(ls(pattern = "sim_single_brain"), \(sim) {
  obj <- eval(as.symbol(sim))
  reducedDim(obj, "PCAsub") <- NULL
  obj <- as.Seurat(obj, 
                   counts = "counts", 
                   data = "logcounts")
  obj@meta.data <- mutate(obj@meta.data, 
                          perc_deg = paste0(as.numeric(gsub("_.*", "", gsub("sim_single_brain_", "", sim))), "%"), 
                          n_cells = as.character(ncol(obj)), 
                          n_genes = as.character(nrow(obj)), 
                          sce_name = sim)
  return(obj)
})
rm(list = ls(pattern = "sim_single_brain")); gc(full = TRUE)
```

Iterating over the datasets, we print each QC plot & save them to PDFs. 

```{r, fig.width=13, fig.height=8, results='hold', message=FALSE, warning=FALSE}
purrr::walk(obj_list, function(z) {
  # gather metadata 
  obj_name <- z@meta.data$sce_name[1]
  n_cells <- z@meta.data$n_cells[1]
  n_genes <- z@meta.data$n_genes[1]
  perc_deg <- z@meta.data$perc_deg[1]
  # summary stat table 
  sparsity_count <- mean(z@assays$originalexp@counts == 0)
  mean_count <- mean(z@assays$originalexp@counts)
  med_count <- ifelse(sparsity_count > 0.5, 0, median(z@assays$originalexp@counts))
  sd_count <- sd(z@assays$originalexp@counts)
  var_count <- sd_count^2 
  range_count <- range(z@assays$originalexp@counts)
  summary_df <- data.frame(metric = c("Mean", "Median", "S.D.", "Variance", "Range", "Sparsity"), 
                           value = c(round(mean_count, 2), 
                                     round(med_count, 2), 
                                     round(sd_count, 2), 
                                     round(var_count, 2),
                                     paste0("(", range_count[1], ", ", range_count[2], ")"), 
                                     paste0(round(sparsity_count, 4) * 100, "%")))
  # create counts histogram
  p0 <- data.frame(x = as.numeric(z@assays$originalexp@counts)) %>% 
      ggplot(aes(x = x)) + 
      geom_histogram(fill = "deepskyblue3", 
                     color = "dodgerblue4", 
                     alpha = 0.75, 
                     linewidth = 0.75) + 
      scale_y_continuous(labels = scales::label_scientific()) + 
      scale_x_continuous(labels = scales::label_comma()) + 
      labs(x = "Raw Expression", y = "Frequency") + 
      theme_classic(base_size = 14)
  # create log counts histogram
  p1 <- data.frame(x = as.numeric(z@assays$originalexp@data)) %>% 
        ggplot(aes(x = x)) + 
        geom_histogram(fill = "darkseagreen", 
                       color = "forestgreen", 
                       alpha = 0.75, 
                       linewidth = 0.75) + 
        scale_y_continuous(labels = scales::label_scientific()) + 
        labs(x = "Normalized Expression", y = "Frequency") + 
        theme_classic(base_size = 14)
  # create UMAP by cluster 
  p2 <- data.frame(UMAP1 = z@reductions$UMAP@cell.embeddings[, 1], 
                   UMAP2 = z@reductions$UMAP@cell.embeddings[, 2], 
                   cluster = z$label) %>% 
        ggplot(aes(x = UMAP1, y = UMAP2, color = cluster)) + 
        geom_point() + 
        scale_color_manual(values = palette_cluster) + 
        labs(x = "UMAP 1",
             y = "UMAP 2", 
             color = "Louvain Cluster") + 
        theme_classic(base_size = 14) + 
        theme(axis.text = element_blank(), 
              axis.ticks = element_blank()) + 
        guides(color = guide_legend(override.aes = list(size = 4)))
  # create PCA of cell ordering 
  p3 <- data.frame(PC1 = z@reductions$PCA@cell.embeddings[, 1], 
                   PC2 = z@reductions$PCA@cell.embeddings[, 2], 
                   cell_time = z$cell_time_normed) %>% 
        ggplot(aes(x = PC1, y = PC2, color = cell_time)) + 
        geom_point() + 
        scale_color_gradientn(colors = paletteer_d("wesanderson::Zissou1")) + 
        labs(x = "PC 1", 
             y = "PC 2", 
             color = "True Ordering") + 
        theme_classic(base_size = 14) + 
        theme(axis.text = element_blank(), 
              axis.ticks = element_blank())
  # table of simulation parameters 
  param_df <- data.frame(metric = c("Number of Cells", "Number of Genes", "% Dynamic Genes"), 
                         value = c(as.character(n_cells), as.character(n_genes), perc_deg))
  plot_table <- rbind(summary_df, param_df) %>% 
                gridExtra::tableGrob(rows = NULL, 
                                     cols = c("Metric", "Value"), 
                                     theme = gridExtra::ttheme_minimal(core = list(fg_params = list(hjust = 0, x = 0.05)), 
                                                                       colhead = list(fg_params = list(hjust = 0, x = 0.05)))) %>% 
                gtable::gtable_add_grob(grobs = grid::rectGrob(gp = grid::gpar(fill = NA, lwd = 2)), 
                                        t = 2, 
                                        b = nrow(.), 
                                        l = 1, 
                                        r = ncol(.)) %>% 
                gtable::gtable_add_grob(grobs = grid::rectGrob(gp = grid::gpar(fill = NA, lwd = 2)), 
                                        t = 1, 
                                        l = 1, 
                                        r = ncol(.))
  # align everything 
  p4a <- (p0 | p1) / (p2 | p3) + 
         plot_layout(guides = "collect")
  p4b <- (p4a | plot_table) + 
         plot_layout(ncol = 2, widths = c(3, 1)) + 
         plot_annotation(title = paste0("Metrics for dataset: ", obj_name), 
                         theme = theme(plot.title = element_text(face = "bold", size = 16)))
  # save & print plot
  ggsave(filename = paste0("QC_", obj_name, ".pdf"),
         plot = p4b, 
         device = "pdf", 
         path = "/blue/rbacher/j.leary/repos/scLANE-Simulations/Figures/Sim_QC/", 
         width = 13,
         height = 8, 
         units = "in", 
         dpi = "retina")
  print(p4b)
  # cleanup 
  sink(tempfile())
  rm(p0, p1, p2, p3, p4a, p4b, plot_table); gc(full = TRUE)
  sink()
})
rm(obj_list)
```











# Session Info 

```{r}
sessioninfo::session_info()
```
